---
title: "Handball WC23 Analysis"
author: "Andreas Cardeneo"
format: html
editor: visual
---

## Setup

We first load the required libraries

```{r}
library(rvest)
library(tidyverse)
library(furrr)
```

### Web resources

Next, we set the *International Handball Federation* (IHF) URLs where we will retrieve the data from.

```{r}
ihf_base <- "https://www.ihf.info"
ihf_wc2023 <- "/competitions/men/308/28th-ihf-mens-world-championship-2023/101253"
ihfurl <- paste0(ihf_base, ihf_wc2023)
```

## Loading data

With these setup steps, we load the start page where we will find the links to each team.

```{r}
page <- read_html(ihfurl)
```

## Process pages

To process the HTML pages we define two small functions that will help to make our analysis script a little bit clearer syntactically. Note that these functions contain the selectors used to identify the correct places in the HTML documents where the team overview resp. the player information can be found. These selectors have been found using the developer tools of Google Chrome.

```{r}
team_list <- function(src) {
  src |> html_element(css = "#block-views-block-teams-flags-list-block-1-2 > div > div")
}

process_player <- function(player_url, base_url) {
  player_page <- read_html(paste0(base_url, player_url))
  tmp <- player_page |> html_elements("#block-ihf-theme-content > div > div > div > div > div.col-md-7") |>
    html_nodes("h3")
  elems <- tmp |> html_text2() 
  tibble(BIRTHDAY = elems[[3]], AGE = elems[[5]], 
         HEIGHT = str_sub(elems[[6]], 1, 3) |> as.integer(), 
         WEIGHT = str_sub(elems[[7]], 1, -3) |> as.integer())
}

process_team <- function(team_url, base_url) {
  team_page <- read_html(paste0(base_url, team_url))
  tmp <- team_page |> html_elements("#block-ihf-theme-content > div > div > section > div > div:nth-child(1) > div > div.row") |>
    html_nodes("a") 
  data <- tmp |> html_text2()
  players <- str_split_fixed(data, "\n", n = 4) |> as_tibble(.name_repair = "unique")
  names(players) <- c("FIRST_NAME", "SECOND_NAME", "CLUB", "POSITION")
  players$HREF <- tmp |> html_attr("href")
  
  details <- players$HREF |> map_dfr(process_player, base_url = base_url, .progress = TRUE)
  bind_cols(players, details)
}
```

### Team overview page

```{r}
team_urls <- page |> team_list() |> html_nodes("a") |> html_attr("href")

thelist <- team_urls

```

The variable `thelist` is introduced to simplify debugging and testing as it is easy to reduce the size of the team list.

### Player data

We process each link to a team with the above defined `process_team` function that in turn loads each player profile from the player's page. We do so in parallel using some worker processes using the `furrr` package.

```{r}
plan(multisession, workers = 8)

player_table <- thelist |> 
  future_map_dfr(process_team, base_url = ihf_base, .progress = TRUE) 


```

Having done that, we need to process the strings we find in the extracted text elements in order to extract (and clean) team names, nations and positions.

```{r}
cntry_pos <- player_table$POSITION |> 
  str_match("(?<COUNTRY>.+)\\s-\\s(?<POSITION>.+)") |> 
  as_tibble(.name_repair="unique")
player_table$COUNTRY <- cntry_pos$COUNTRY
player_table$POSITION <- cntry_pos$POSITION
clb_name <- player_table$CLUB |> 
  str_match("Club:\\s(?<NAME>.+)") |> 
  as_tibble(.name_repair = "unique")
player_table$CLUB <- clb_name$NAME
```

Now we safe the dataset locally as it takes quite some time to retrieve and build it.

```{r}
player_table |> write_csv2("./player_table_wc23.csv")
```

## Corrections

Still missing: There are errors in the data on the website. These need to be corrected manually.

Some player pages, especially for Chile, are wrongly formatted and height cannot be read by the method above.

## Analysis

### Teams that send the most players to the championship

```{r}
player_table |> 
  group_by(CLUB) |> 
  summarise(N = n()) |> 
  slice_max(n = 15, order_by = N) |>
  ggplot(aes(x = reorder(CLUB, N), y = N)) + geom_col(fill = "blue") + 
  coord_flip() + xlab("Team") + ylab("Number of players")
```

### Number of different teams national teams come from

```{r}
player_table |> 
  group_by(COUNTRY) |> 
  summarise(NT = n_distinct(CLUB)) |> 
  slice_min(n = 5, order_by = NT) |>
  ggplot(aes(x = reorder(COUNTRY, -NT), y = NT)) + geom_col(fill = "blue") +
  coord_flip() + xlab("National team") + 
  ylab("Number of different teams players come from")
```
